Supporting Information:\\ Accurate Binding of Sodium and Calcium to a POPC Bilayer by Effective Inclusion of Electronic Polarization

### 2018-03-12 Manuscript/SI.tex
Updates based on reviewers' comments

\author{Josef Melcr}
\author{Hector Martinez-Seara}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\author{Ji{\v r}{\' i} Kolafa}
\affiliation{Department of Physical Chemistry, Institute of Chemical Technology, Prague 6, Czech Republic}
\author{Pavel Jungwirth}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\alsoaffiliation{Department of Physics, Tampere University of Technology, P.O. Box 692, FI-33101
Tampere, Finland}

@@

\title{Supporting Information:\\Accurate binding of sodium and calcium to phospholipid bilayers by effective inclusion of electronic polarization}

@@

>>>>

\author{Josef Melcr}
\author{Hector Martinez-Seara}
\author{Ricky Nencini}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\author{Ji{\v r}{\' i} Kolafa}
\affiliation{Department of Physical Chemistry, Institute of Chemical Technology, Prague 6, Czech Republic}
\author{Pavel Jungwirth}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\alsoaffiliation{Department of Physics, Tampere University of Technology, P.O. Box 692, FI-33101
Tampere, Finland}

@@

\title{Supporting Information:\\ Accurate Binding of Sodium and Calcium to a POPC Bilayer by Effective Inclusion of Electronic Polarization}

@@

\newpage
\section{Electrostatic potential of lipid bilayer}
\begin{figure}[!h]
  \centering
  \includegraphics[width=16.0cm]{../Fig/potential.eps}
  \caption{\label{potential}
    Electrostatic potential calculated by evaluating the double integral of the charge density
    using {\it gmx potential} tool from Gromacs package \cite{gromacsMANUAL}.
    Comparing the Lipid14 and ECC-POPC simulations with TIP3P water model
    reveals that applying ECC to the headgroup and glycerol backbone charges slightly decreases
    the dipole potential inside the bilayer. On the other hand, ECC-POPC simulation
    with SPC/E water model has larger dipole pontential than with TIP3P.
    The stronger dependence of water model is expected as the dipole potential
    mainly originates from orientation of water molecules \cite{gawrisch92}.
  }
\end{figure}

\newpage
\section{Binding of potassium to ECC-POPC lipid bilayer}
\begin{figure}[!h]
  \centering
  \includegraphics[width=13.0cm]{../Fig/density_profiles_KCl-water.eps}
  \caption{\label{Kdens}
    Number density profiles of K$^+$, Cl$^-$ and water along membrane normal
    axis from ECC-POPC simulation with KCl ($C_{ion}'$=1000mM, simulation parameters as in Table S1).
    Parameters for ions are taken from Ref. \citenum{mason12}.
    In consistent with figure 5 in the main text, the density profiles of ions and water
    are divided with 2 and 200, respectively. Similarity of potassium
    and chloride curves together with the bulk concentrations indicate lower binding 
    affinity for potassium than for sodium (Fig. 5 in the main text).
    Since this is in line with wide range of experimental data \cite{eisenberg79,binder02,klasczyk10}, we conclude
    that the ECC-POPC model reproduces the correct binding affinity also for potassium.
    In line with the low binding affinity, the headgroup order parameters were not
    affected by the presense of KCl in simulations. The order parameter 
    data is not shown because we are not aware of experimental data for PC headgroup
    order parameters as a function of KCl concentration.
  }
\end{figure}
